---
output: html_document
---
Peer Graded Assignment: Course Project 1
========================

#####Author: Jonathan Sinclair

load relevant packages & data
```{r results='hide', message=FALSE, warning=FALSE}
library(dplyr)
##Assumes Zip folder is in working directory (still zipped)
act <- read.csv(unz("repdata_data_activity.zip", "activity.csv"))
```

###What is the mean total number of steps taken per day

**1) Calculate the total number of steps taken per day**

```{r}
## Convert act$date to as.Date()
act$date <- as.Date(x = act$date, "%Y-%m-%d")

steps_per_day <- act %>% 
  group_by(date) %>% 
  summarize(steps=sum(steps, na.rm=TRUE))

total_steps_per_day <- sum(steps_per_day$steps)
print(paste("Total number of steps taken per day =", total_steps_per_day))
```

**2) Make a Histogram of the total number of steps taken each day**

```{r}
hist(steps_per_day$steps, main="Histogram: Total number of steps per day", ylab="Count", xlab="Total number of steps per day")
```

**3) Average and Median steps per day**

```{r}
      ## Mean
mean_steps_per_day <- mean(steps_per_day$steps)
print(paste("Average steps taken per day =", mean_steps_per_day))
      ## Median
median_steps_per_day <- median(steps_per_day$steps)
print(paste("Median steps taken per day =", median_steps_per_day))
```

...................................

##What is the average daily activity pattern?

**1) Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days**

```{r}
# Proccess / Transform data
average_steps_per_day <- act %>%
  group_by(interval) %>%
  summarize(av_steps=mean(steps, na.rm=TRUE))

# Make Plot
plot(average_steps_per_day, type="l", main="Average number of steps taken across intervals", xlab="Interval", ylab="Average steps")
```


**2) Which 5 minute interval on average, contains the maximum number of steps?**

```{r}
which.max(average_steps_per_day$av_steps)
```
Interval 104








